### 2015-09-30 referee_report/response_to_referee.txt
Clean up the repo structure by making referee report directory, which contains the .doc file.

From: Dr. Michael Gully-Santiago, on behalf of all coauthors
To: Anonymous referee and topical editor, Applied Optics
Date: Monday, September 21 - Friday September 25, 2015
Re: Response to referee report

Thank you for your referee report and detailed reading of our manuscript, "Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy".  Below I respond to the referee's feedback.

I corrected the accent mark in Fabry-Pérot for all instances of the name.

I corrected these grammatical errors:
- p5. Since silicon is opaque in the optical (range?), the...
- p6. are too small to (be) resolve(d) spatially

I removed the sentence "Our ultimate goal is to bond optics with thicknesses as large as 30 mm.", which we agree is too specific to our application.

I reduced the abstract length by about 40% to make it more concise.

I corrected the title of Section 3.

I entirely removed the paragraph in Section 4, "The Cary 5000 has four operation modes..."

I entirely removed the paragraph in Section 4, "The spectral sampling ranged from..."  (This uncontroversial information is in the nearby table.)

I added a discussion of least squares regression versus Gaussian process regression.

In regards to the comment, "For me Tdsp is missing in eq 5. ( Tmix= f Tg&#43;(1-f)Tdsp )", the referee is correct that the last term in eq. 5 represents the contribution of DSP wafer, however the Tdsp factor has been divided out as described in eq 4.  So the last term could be thought of as (1-f)*Tdsp/Tdsp = (1-f)*1.  The reviewers comment did, however, cause me to revise my description of equation 4 to clarify this point.

In regards to the comment, '5.B: the term "Sample Transport Accessory" is really not adapted...', I have removed the mention of the sample transport accessory, and instead simply described the process of stepping across the sample.

I have now defined temperature, T, before its first use, and have eliminated the Figure 1, which was a plot of the temperature dependence of the refractive index.

In response to the referee's comments on concision, I have cut many lines of text, and in some cases entire paragraphs that I agreed were immaterial to the main idea of the paper.  The manuscript lost at least 2 pages out of the raw 32.  I have motivated more clearly the interest in the MCMC formalism over a conventional least squares technique, now located in section 4- "comparing models to data."  Indeed, I agree that a least squares technique can be applied, but I also note the advantages of the Gaussian process regression technique.  I have posted several references to help the reader reproduce the Gaussian Process regression technique if desired.

Upon inspecting this final submission, it is clear that the manuscript has been much improved from the reviewer's comments, particularly with a clearer 'flow' to the paper.

Best regards,
Dr. Gully-Santiago